Modelling the Retinotectal projection - can competition provide a stopping mechanism?

### 2022-02-23 paper/paper.tex
Paper WIP

% Use the article doc class, with an 11 pt basic font size
%\documentclass[11pt, a4paper, draft]{article}
\documentclass[11pt, a4paper]{article}

@@

The model defined by the receptor and ligand expressions of Eqs.\,\ref{e:retrcpt}
\& \ref{e:tecligexp} along with Eq.\,\ref{e:mv2} has three
free parameters; $m_{\!_G}$, $m_c$ and $r_c$. To find the best combination of
these parameters, we reasoned that evolution would optimize the arrangement of
the wildtype pattern. We performed a simulated annealing optimization to
minimize the error metric, $\epsilon$, for the wildtype pattern
(Fig.\,\ref{f:G}E) at simulation time 1000. The parameters returned were
$m_{\!_G}$ = 0.003841, $m_c$ = 0.09959 and $r_c$ = 0.39918. Fig.\,\ref{f:GC}
shows the $\mathbf{GC}$ simulation with these parameters. Fish net plots at
$t=40$ and $t=1000$ show that the pattern develops quickly and compares well
with the expected layout (Fig.\,\ref{f:G}E). We investigated the response of
this model to the same set of surgical manipulations shown for the
chemoaffinity-only model in Fig.\,\ref{f:Gsurg}. We found that the addition of
competition leads to a better correspondence to the experimental results than
for the chemoaffinity-only model (see supplementary Fig.\,\ref{f:GCsurg}).

@@

% First, look at I, from Simpson and Goodhill.
First, we took inspiration from Simpson and Goodhill's `I' interaction, which
is defined as:

%We modelled interactions between axon growth cones due to signalling between
%receptors of the same type. For these interactions, repulsion caused a
%movement of branch $b$ along a unit vector, $\hat{\mathbf{x}}_{kb}$, from
%branch $k$ to branch $b$ causing them to move further apart; attraction caused
%the opposite movement. This interaction, $\mathbf{I}_b$, acting on branch $b$,
%is given by
%
\begin{equation}
\mathbf{I}_b
= \frac{1}{|B_{b,I}|} \sum_k^{nM} \sum_i^N \hat{\mathbf{x}}_{kb}\,W_{\!_I} \qquad \mathrm{if}~r_{i,b}
/ r_{i,k} > 1.1
\end{equation}
%
the weight here is defined as a function of $r_{\!_I}$:
%
\begin{equation}
W_{\!_I}(d_{kb}) = \begin{cases}
      1 - \frac{d_{kb}}{2r_{\!_I}}   & d_{kb} \leq 2r_{\!_I} \\
     0 & d_{kb} > 2r_{\!_I}
     \end{cases}
\end{equation}
$r_{i,k}$ is the expression of receptor $i$ on branch $k$.

 (supp?) Fig.\,\ref{f:GI} and Fig.\,\ref{f:GIsurg}.

@@

>>>>

% Use the article doc class, with an 11 pt basic font size
\documentclass[11pt, a4paper, draft]{article}
%\documentclass[11pt, a4paper]{article}

@@

The model defined by the receptor and ligand expressions of Eqs.\,\ref{e:retrcpt}
\& \ref{e:tecligexp} along with Eq.\,\ref{e:mv2} has three
free parameters; $m_{\!_G}$, $m_c$ and $r_c$. To find the best combination of
these parameters, we reasoned that evolution would optimize the arrangement of
the wildtype pattern. We performed a simulated annealing optimization to
minimize the error metric, $\epsilon$, for the wildtype pattern
(Fig.\,\ref{f:G}E) at simulation time 1000. The parameters returned were
$m_{\!_G}$ = 0.003841, $m_c$ = 0.09959 and $r_c$ = 0.39918. Fig.\,\ref{f:GC}
shows the $\mathbf{GC}$ simulation with these parameters. Fish net plots at
$t=40$ and $t=1000$ show that the pattern develops quickly and compares well
with the expected layout (Fig.\,\ref{f:G}E). We investigated the response of
this model to the same set of surgical manipulations shown for the
chemoaffinity-only model in Fig.\,\ref{f:Gsurg}. We confirmed that the
addition of competition leads to a better correspondence to the experimental
results than for the chemoaffinity-only model (see supplementary
Fig.\,\ref{f:GCsurg}).

@@

% First, look at I, from Simpson and Goodhill.
First, we took inspiration from Simpson and Goodhill's model for the
interaction between EphA receptors, which is defined (with receptor type $i$
standing for EphA) as:

%We modelled interactions between axon growth cones due to signalling between
%receptors of the same type. For these interactions, repulsion caused a
%movement of branch $b$ along a unit vector, $\hat{\mathbf{x}}_{kb}$, from
%branch $k$ to branch $b$ causing them to move further apart; attraction caused
%the opposite movement. This interaction, $\mathbf{I}_b$, acting on branch $b$,
%is given by
%
\begin{equation}
\mathbf{I}_b
= \frac{1}{|B_{b,I}|} \sum_k^{nM} \hat{\mathbf{x}}_{kb}\,W_{\!_I} \qquad \mathrm{if}~r_{i,b}
/ r_{i,k} > 1.1
\end{equation}
%
$r_{i,k}$ is the expression of receptor $i$ on branch $k$ and the weight here
is defined as a function of the distance, $d_{kb}$, between branches:
%
\begin{equation}
W_{\!_I}(d_{kb}) = \begin{cases}
      1 - \frac{d_{kb}}{2r_{\!_I}}   & d_{kb} \leq 2r_{\!_I} \\
     0 & d_{kb} > 2r_{\!_I}
     \end{cases}
\end{equation}
%
$B_{b,I}$ is the set of growth cones within a distance $2r_{\!_I}$ of branch
$b$. $\hat{\mathbf{x}}_{kb}$ is as defined in Eq.\,\ref{e:C}.

Simpson and Goodhill introduced this interaction as an additive modification
to the main, distance based competition mechanism of Eq.\,\ref{e:C} (adding a
term $m_{\!_I} \mathbf{I}_b$ to the equation). They introduced an interaction
varying along the temporal-nasal retinal axis and mediated by one receptor,
EphA, for which there was experimental data supporting a repulsive
effect. They noted that the use of a receptor ratio (comparing the ratio with
a `discrimination limit', above which repulsion would be activated) is
equivalent to a multiplicative interaction between EphA receptors and ephrin-A
ligands with the strong assumption that ligand expression is the inverse of
receptor expression, i.e. that $l_i = 1/r_i$.

We asked i) whether this mechanism could \emph{replace} the non-biological
competition, C, (making necessary its extension to the dorsal-ventral retinal
axis) and ii) if cast in terms of receptor-ligand interactions, could the
strong ligand expression assumption be relaxed?

% (supp?) Fig.\,\ref{f:GI} and Fig.\,\ref{f:GIsurg}.

@@

In order to replace the distance-based competition, $\mathbf{C}_b$, we first
suggest that each receptor ligand pair involved in the retinotectal system
could mediate a repulsive interaction. DO WE NEED TO? Taken alone, the EphA
interaction between one branch, $b$, and another, $k$, will cause a repulsion
if $k$ originated more \textbf{nasally?} than $b$. What about a branch
originating on the temporal side of $b$? This could be handled by the
interaction between $k$ and $b$ and may not require an additional interaction.
